*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-07 Code/Manuscript_O2.Rmd
Added OPP dataset citations

*Prochlorococcus marinus*, the smallest picocyanobacterium, comprises multiple clades with distinct niches across tropical and sub-tropical oligotrophic ocean regions, including Oxygen Minimum Zones.  Ocean warming may open permissive temperatures in  new, poleward photic regimes, along with expanded Oxygen Minimum Zones.  We used ocean protein data [CITATION] to help guide testing of *Prochlorococcus marinus* growth across a matrix of peak irradiances, photoperiods, spectral bands and dissolved oxygen.  MED4 from Clade HLI requires greater than 4 h photoperiod, grows at 25 µmol O~2~ L^-1^ and above, and exploits high cumulative diel photon doses, yet shows accelerated growth when the cost of photoinactivation is lowered under red, vs. blue, actinic light.  MED4 relies upon an alternative oxidase to balance electron transport, which may exclude it from growth under our lowest 2.5 µmol O~2~ L^-1^ condition.  SS120 from Clade LLIII is restricted to low light under full 250 µmol O~2~ L^-1^ , shows expanded light exploitation under  25 µmol O~2~ L^-1^, but is excluded from growth under  2.5 µmol O~2~ L^-1^.  Intermediate oxygen suppresses the cost of PSII photoinactivation, and enzymatic production of H~2~O~2~ in SS120, which has limited genomic capacity for PSII and DNA repair.  MIT9313 from Clade LLIV is restricted to low blue irradiance under 250 µmol O~2~ L^-1^, but exploits much higher irradiance under red light, or under lower [O~2~], conditions which slow photoinactivation of PSII and production of reactive oxygen species.

>>>>

*Prochlorococcus marinus*, the smallest picocyanobacterium, comprises multiple clades with distinct niches across tropical and sub-tropical oligotrophic ocean regions, including Oxygen Minimum Zones.  Ocean warming may open permissive temperatures in new, poleward photic regimes, along with expanded Oxygen Minimum Zones.  We used ocean protein data [@saitoPeptidesTheirSpectral2018; @saitoDevelopmentOceanProtein2021] to help guide testing of *Prochlorococcus marinus* growth across a matrix of peak irradiances, photoperiods, spectral bands and dissolved oxygen.  MED4 from Clade HLI requires greater than 4 h photoperiod, grows at 25 µmol O~2~ L^-1^ and above, and exploits high cumulative diel photon doses, yet shows accelerated growth when the cost of photoinactivation is lowered under red, vs. blue, actinic light.  MED4 relies upon an alternative oxidase to balance electron transport, which may exclude it from growth under our lowest 2.5 µmol O~2~ L^-1^ condition.  SS120 from Clade LLIII is restricted to low light under full 250 µmol O~2~ L^-1^ , shows expanded light exploitation under  25 µmol O~2~ L^-1^, but is excluded from growth under 2.5 µmol O~2~ L^-1^.  Intermediate oxygen suppresses the cost of PSII photoinactivation, and enzymatic production of H~2~O~2~ in SS120, which has limited genomic capacity for PSII and DNA repair.  MIT9313 from Clade LLIV is restricted to low blue irradiance under 250 µmol O~2~ L^-1^, but exploits much higher irradiance under red light, or under lower O~2~ concentrations, conditions which slow photoinactivation of PSII and production of reactive oxygen species.
